---
title: "Mouse vs Human"
author: "Ronica K"
output: 
  html_document:
    keep_md: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=8, dpi=300, echo=F, warning=F, message=F, cache=T)
```


```{r}
library(BSgenome.Hsapiens.UCSC.hg19)
library(BSgenome.Mmusculus.UCSC.mm10)
library(ggplot2); library(ggbio); library(RColorBrewer) 
library(reshape2); library(MASS); library(e1071)
```

###Mouse genome

__UCSC RepeatMasker Track__

```{r}
source('../txtToGR.R')

#Human genome
hgnames <- paste0('chr', c(1:22, 'X', 'Y'))
hgseqlen <- as.numeric(seqlengths(BSgenome.Hsapiens.UCSC.hg19)
                       [seqnames(BSgenome.Hsapiens.UCSC.hg19) %in% hgnames])
hg19gr <- wrapper('../GenomeTables/hg19rmsk.txt.gz', 'hg19', hgnames, hgseqlen)

#Mouse genome
mmnames <- paste0('chr', c(1:19, 'X', 'Y'))
mmseqlen <- as.numeric(seqlengths(BSgenome.Mmusculus.UCSC.mm10)
                       [seqnames(BSgenome.Mmusculus.UCSC.mm10) %in% mmnames])
mm10gr <- wrapper('../GenomeTables/mm10rmsk.txt.gz', 'mm10', mmnames, mmseqlen)


mm10gr
```

__Number of unique repeat sequences__

```{r}
length(unique(mm10gr$name))
```

```{r}
classColourCount <- length(unique(mm10gr$class))
```

###Repeats

```{r}
hg19classLength <- data.frame(chr=seqnames(hg19gr), length=width(hg19gr), 
                          class=hg19gr$class, family=hg19gr$family, genome='hg19')
mm10classLength <- data.frame(chr=seqnames(mm10gr), length=width(mm10gr), 
                          class=mm10gr$class, family=mm10gr$family, genome='mm10')

classLength <- rbind(hg19classLength, mm10classLength)

summary(hg19classLength)
summary(mm10classLength)
```

__Number of repeats by class__

```{r}
numrep <- sapply(split(classLength, classLength$genome), 
                 function(x) sapply(split(x, x$class), function(y) nrow(y)))
numrep
```

__Genome covered by class__

```{r}
covrep <- sapply(split(classLength, classLength$genome), 
                 function(x) sapply(split(x, x$class), function(y) sum(y$length)))
covrep
```

__Length by species__

```{r fig.height=3.5}
ggplot(classLength, aes(genome, log(length), fill=genome)) + geom_boxplot() +
    theme(legend.position='none') + coord_flip() + scale_fill_brewer(palette='Dark2')
```


__Length by class__

__in mouse__

```{r}
sapply(split(mm10classLength, mm10classLength$class), function(x) summary(x$length))
```


```{r}
ggplot(classLength, aes(genome, log(length))) + 
  geom_jitter(aes(colour=genome), alpha=0.05, position=position_jitter(width = .35)) + 
  geom_boxplot(alpha=0) + scale_y_continuous(name='log(length)') + 
  theme(legend.position='none') + scale_color_brewer(palette='Dark2') +
  facet_grid(.~class)
```

###Gaps

```{r}
hg19NoStrand <- hg19gr
strand(hg19NoStrand) <- '*'

mm10NoStrand <- mm10gr
strand(mm10NoStrand) <- '*'

hg19GapByClassLength <- do.call(
  'rbind', lapply(split(hg19NoStrand, hg19NoStrand$class), function(x){
    class <- unique(x$class)[1]
    starts <- min(start(split(x, seqnames(x))))
    ends <- max(end(split(x, seqnames(x))))
    gr <- gaps(x, start=starts, end=ends)
    gr <- gr[strand(gr)=='*']
    data.frame(chr=seqnames(gr), length=width(gr), class=class, genome='hg19')
  }))

mm10GapByClassLength <- do.call(
  'rbind', lapply(split(mm10NoStrand, mm10NoStrand$class), function(x){
    class <- unique(x$class)[1]
    starts <- min(start(split(x, seqnames(x))))
    ends <- max(end(split(x, seqnames(x))))
    gr <- gaps(x, start=starts, end=ends)
    gr <- gr[strand(gr)=='*']
    data.frame(chr=seqnames(gr), length=width(gr), class=class, genome='mm10')
  }))

gapByClassLength <- rbind(hg19GapByClassLength, mm10GapByClassLength)

summary(hg19GapByClassLength)
summary(mm10GapByClassLength)
```

__Gap length by species__

```{r fig.height=3.5}
ggplot(gapByClassLength, aes(genome, log(length), fill=genome)) + geom_boxplot() +
    theme(legend.position='none') + coord_flip() + scale_fill_brewer(palette='Dark2')
```

__Gap length by class__

__in mouse__

```{r}
sapply(split(mm10GapByClassLength, mm10GapByClassLength$class), function(x) summary(x$length))
```

```{r}
ggplot(gapByClassLength, aes(genome, log(length))) + 
  geom_jitter(aes(colour=genome), alpha=0.05, position=position_jitter(width = .35)) + 
  geom_boxplot(alpha=0) + scale_y_continuous(name='log(length)') + 
  theme(legend.position='none') + scale_color_brewer(palette='Dark2') +
  facet_grid(.~class)
```


### Frequency

__Number of repeats by sequence name__

```{r}
#number of repeats by sequence
seqnum <- table(mm10gr$name)
seqfreq <- do.call(rbind, Map(data.frame, num.ranges=seqnum))
head(seqfreq, 20)
```

__Frequency distribution__

```{r}
summary(seqfreq)
```

__Square root scale, red median, blue 1000 reps per sequence__

```{r}
ggplot(seqfreq, aes(num.ranges)) + geom_histogram(binwidth=5, colour='white') +
  geom_vline(xintercept=median(seqnum), color='red') +  
  geom_vline(xintercept=1000, color='blue') +
  scale_x_continuous(trans='sqrt')  
```

__Over 1000 reps per sequence__

```{r}
seqfreq$seq <- row.names(seqfreq)
over1000reps <- seqfreq$seq[seqfreq$num.ranges>1000]
length(over1000reps)
over1000reps
```

__Over 1000 subset__

```{r}
over1000subset <- mm10gr[elementMetadata(mm10gr)[,1] %in% over1000reps]
over1000subset
```

__Ratio__

```{r}
length(over1000subset)/length(mm10gr)
```

__Length by class__

```{r}
over1000classLength <- data.frame(chr=seqnames(over1000subset),
                                  length=width(over1000subset), 
                                  class=over1000subset$class, 
                                  family=over1000subset$family)
sapply(split(over1000classLength, over1000classLength$class), function(x) summary(x$length))
```

__Length ranges__

```{r}
over1000classLength$range <- ifelse(over1000classLength$length<=50, '<=50', '>50')
table(over1000classLength$range)

```

__Shorter than or equal to 50 bases__

```{r}
summary(over1000classLength[over1000classLength$range=='<=50',])

```

__Longer than 50 bases__

```{r}
summary(over1000classLength[over1000classLength$range=='>50',])

```