---
title: "Coverage and reduced intervals"
author: "Ronica K"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=F, message=F, warning=F}
library(GenomicRanges)
library(ggplot2)


plotRanges <- function(x, xlim=x, main=deparse(substitute(x)),
                       col='black', sep=0.5,...){
  height <- 1
  if (is(xlim, 'Ranges'))
    xlim <- c(min(start(xlim)), max(end(xlim)))
  bins <- disjointBins(IRanges(start(x), end(x) + 1))
  plot.new()
  plot.window(xlim, c(0, max(bins)*(height+sep)))
  ybottom <- bins*(sep+height)-height
  rect(start(x)-0.5, ybottom, end(x)+0.5, ybottom+height, col=col,...)
  title(main)
  axis(1)
}
```

###Coverage and Rle

Some NON-overlapping ranges on chromosome 1 (30 bp length) of some NON-existing creature

```{r, echo=F, cache=T}
r <- GRanges(seqnames = "chr1", strand = c("+", "+", "+", "+"), ranges = IRanges(start = c(6,13,18,24), width = 3))
seqlengths(r) <- 30
r
```

```{r, echo=F, fig.height=2.5}
plotRanges(ranges(r))
```

- Coverage 

Lengths - length of a genome segment 

Values - number of ranges that cover the segment

```{r, echo=F}
rcov <- coverage(r)
rcov
```


This one is a direct representation of coverage

Chromosome length on x-axis, number of covering ranges on y-axis

```{r, echo=F}
plot(as.numeric(rcov$chr1))
```

- Add some overlap (ranges 2 and 3)

```{r, echo=F, cache=T}
rO <- GRanges(seqnames = "chr1", strand = c("+", "+", "+", "+"), ranges = IRanges(start = c(6,13,18,24), width = c(3,6,3,3)))
seqlengths(rO) <- 30
rO
```

```{r, echo=F, fig.height=3.3}
plotRanges(ranges(rO))
```

Note the difference in Lengths and Values 

```{r, echo=F}
rOcov <- coverage(rO)
rOcov
```

```{r, echo=F}
plot(as.numeric(rOcov$chr1))
```


- And some more  

```{r, echo=F, cache=T}
r2O <- GRanges(seqnames = "chr1", strand = c("+", "+", "+", "+"), ranges = IRanges(start = c(6,13,18,24), width = c(9,6,8,3)))
seqlengths(r2O) <- 30
r2O
```

```{r, echo=F, fig.height=3.3}
plotRanges(ranges(r2O))
```

```{r, echo=F}
r2Ocov <- coverage(r2O)
r2Ocov
```

```{r, echo=F}
plot(as.numeric(r2Ocov$chr1))
```


- And even more  

```{r, echo=F, cache=T}
r3O <- GRanges(seqnames = "chr1", strand = c("+", "+", "+", "+"), ranges = IRanges(start = c(6,13,18,24), width = c(13,8,8,3)))
seqlengths(r3O) <- 30
r3O
```

```{r, echo=F, fig.height=4}
plotRanges(ranges(r3O))
```

```{r, echo=F}
r3Ocov <- coverage(r3O)
r3Ocov
```

```{r, echo=F}
plot(as.numeric(r3Ocov$chr1))
```


###Reduced ranges

Get back to the first set of ranges with no overlaps

There we have gaps of lengths 2, 3 and 4 bp
```{r, echo=F}
r
```

```{r, echo=F}
coverage(r)
```

```{r, echo=F}
plot(c(0,30), c(0,0), type="n", xlab=seqnames(r)[1], ylab="")
arrows(start(r),rep(0,length(r)),
       end(r),rep(0,length(r)),
       lwd=3, length=.1)
```

Lets reduce two ranges into one if they are 

- 2 (or less) bp apart

```{r, echo=F}
red2 <- reduce(r, min.gapwidth=2+1)
red2
```

```{r, echo=F}
coverage(red2)
```

```{r, echo=F}
plot(c(0,30), c(0,0), type="n", xlab=seqnames(red2)[1], ylab="")
arrows(start(red2),rep(0,length(red2)),
       end(red2),rep(0,length(red2)),
       lwd=3, length=.1)
```

- 3 or less

```{r, echo=F}
red3 <- reduce(r, min.gapwidth=3+1)
red3
```

```{r, echo=F}
coverage(red3)
```

```{r, echo=F}
plot(c(0,30), c(0,0), type="n", xlab=seqnames(red3)[1], ylab="")
arrows(start(red3),rep(0,length(red3)),
       end(red3),rep(0,length(red3)),
       lwd=3, length=.1)
```

- 4 or less

```{r, echo=F}
red4 <- reduce(r, min.gapwidth=4+1)
red4
```

```{r, echo=F}
coverage(red4)
```

```{r, echo=F}
plot(c(0,30), c(0,0), type="n", xlab=seqnames(red4)[1], ylab="")
arrows(start(red4),rep(0,length(red4)),
       end(red4),rep(0,length(red4)),
       lwd=3, length=.1)
```